---
title: HILDA survey
format:
  360-analysis-html:
    # hide code chunk output unless otherwise specified to prevent privacy
    # leaks when previewing data frames
    execute:
      output: false
author: James Goldie
date: last-modified
code-fold: true
---

```{r}
#| label: setup

library(tidyverse)
library(haven)
library(hildareadR)
library(gganimate)
library(here)

hilda_dir <- here("data", "raw-hilda")
```

# Explore the data

Although we'll use the `{hildareadR}` package to combine wave data and extract variables, it's still instructive to open the files using `{haven}` ourselves so that we can look at the column labels and particularly to understand the types of answers available.

This code assumes that all of the files have been unzipped directly into `here("data", "raw-hilda")`. If you've downloaded them from Dataverse and they've unzipped into separate folders by zip file, pull them back up to the parent folder.

```{r}
#| label: load-lastwave
here(hilda_dir, "Combined_v220c.dta") |>
  haven::read_dta() ->
all_lastwave

# expand columns labels out so we can decide which id columns to use
tibble(
  name = names(all_lastwave),
  question = map_chr(all_lastwave, ~ attr(.x, "label")),
  answers = map(all_lastwave, ~ attr(.x, "labels"))) |>
  unnest_longer(answers) |>
  View("Last wave column labels")
```

# Load the data

Variable names in HILDA:

- start with the wave identifier (`a` through `v`)
- end with number from `1` to `20` corresponding to the person the interviewee is referring to (since once person responds on behalf of their household)

```{r}
#| label: vars
survey_vars <- tribble(
  ~ short, ~ long,
  "lsmnhw", "Minutes per week - Housework",
	"lsmncar", "Minutes per week - Caring for disabled spouse/relative",
	"lsmnerr", "Minutes per week - Household errands",
	"lsmnhw", "Minutes per week - Housework",
	"lsmnocd", "Minutes per week - Looking after other children",
	"lshrcar", "Hours per week - Caring for disabled spouse/relative",
	"lshrchd", "Hours per week - Playing with your children",
	"lshrerr", "Hours per week - Household errands",
    #  'such as shopping, banking, paying bills, and keeping financial records (but do not include driving children to school and to other activities)'
	"lshrhw", "Hours per week - Housework",
    # 'such as preparing meals, washing dishes, cleaning house, washing clothes, ironing and sewing'
	"lshrocd", "Hours per week - Looking after other children",
	"lshrod", "Hours per week - Outdoor tasks",
    # 'including home maintenance (repairs, improvements, painting etc.), car maintenance or repairs and gardening'
  "lscar", "Combined - Caring for disabled spouse/relative",
  "lschd", "Combined - Playing with your children",
  "lscom", "Combined - Travelling to/from paid employment",
  "lsemp", "Combined - Paid employment",
  "lserr", "Combined - Household errands",
  "lshw", "Combined - Housework",
  "lsocd", "Combined - Playing with other peoples children",
  "lsod", "Combined - Outdoor tasks",
  "lsvol", "Combined - Volunteer/Charity work",
)

all_lastwave |>
  select(ends_with(survey_vars$short)) ->
timeobs

attr(all_lastwave$vhhstate, "labels")
```

The `{hildareadR}` package saves us a lot of the work of combining waves and selecting out variables of interest, as it handles the wave prefixes (although it is instructive to open the files in `{haven}` manually).

The variables of interest around hours/minutes per week spent on various activities all relate to the _reporting_ person's file. So we don't need `hgageNN` and `hgsexNN`, which relate to other people in the household — we just need `hgage` and `hgsex`, which related to the person doing the reporting.

```{r}
#| label: load-all
read_hilda(
  c("hgsex", "hgage", variable_suffixes),
  waves = 1:22, dir = hilda_dir, release = 22) |>
  select(
    -contains(paste0("hgage", 1:20)),
    -contains(paste0("hgsex", 1:20))) |>
  select(xwaveid, wave, hgage, hgsex, everything()) ->
all_data
```

Let's first see how much missing data there is. There is formally missing data in R's sense (`NA`), but this survey has also been coded such that negative numbers refer to various error codes (eg. "Did not know"). I'm going to collapse those down as missing too.

```{r}
#| label: allwave-missing
#| output: true
all_data |>
  filter(hgage >= 0, hgsex >= 0) |>
  # replace negative values with NA
  mutate(across(starts_with("ls"), ~ case_when(.x < 0 ~ NA, TRUE ~ .x))) |>
  summarise(across(
    where(is.numeric),
    ~ length(which(is.na(.x))) / length(.x))) |>
  t()
```

Looks like reporting age and sex are all there. For most other variables, around 35–45% of the data are missing.

For `lscarb` and `lscarmn`, nearly 90% are missing. I'm not entirely sure what they are — there's also `lsmncar` (caring minutes) and `lscar` (caring mins/hrs combined), so they must be something else (or perhaps a coding error - they don't appear in the data dictionary). Let's drop them.

For the remaining columns, I'd like to see if the missing data is concentrated in any age/sex/wave group:

```{r}
#| label: group-check-missing
all_data |>
  select(-hgagef, -lscarb, -lscarmn) |>
  filter(hgage >= 0, hgsex >= 0) |>
  # replace negative values with NA
  mutate(across(starts_with("ls"), ~ case_when(.x < 0 ~ NA, TRUE ~ .x))) |>
  mutate(age_group = cut(hgage, c(-Inf, 15, 25, 35, 45, 55, 65, 75, 85, Inf))) |>
  select(xwaveid, wave, hgsex, age_group, hgage, everything()) |>
  group_by(wave, hgsex, age_group) |>
  summarise(across(starts_with("ls"), list(
    n_obs = ~ length(.x),
    pct_miss = ~ length(which(is.na(.x))) / length(.x)))) |>
  ungroup() ->
grouped_missing_stats
```

```{r}
#| label: missing-data-vis
#| output: true
grouped_missing_stats |>
  select(wave, hgsex, age_group, ends_with("pct_miss")) |>
  mutate(group = paste(wave, hgsex, age_group)) |>
  pivot_longer(ends_with("pct_miss"), names_to = "measure") |>
  ggplot() +
    aes(y = group, x = measure, fill = value) +
    geom_tile() +
    scale_fill_distiller(palette = "YlOrRd")
```

Okay, this is too much to look at as one figure. Let's look at it wave by wave:

```{r}
#| label: missing-data-vis-bywave
#| output: true
grouped_missing_stats |>
  select(wave, hgsex, age_group, ends_with("pct_miss")) |>
  mutate(group = paste(wave, hgsex, age_group)) |>
  pivot_longer(ends_with("pct_miss"), names_to = "measure") |>
  filter(wave == "u") |>
  ggplot() +
    aes(y = group, x = measure, fill = value) +
    geom_tile() +
    scale_fill_distiller(palette = "YlOrRd")
```

Much better. If we look through the 20 or so waves by altering the filter, we can see there's a lot of missing data in the 0-15s. That makes sense; I'm kind of surprised there are many of them answering the survey at all. I'll cut the 0-15s.

Group numbers start to tail off in the 75-80 bracket. They're not small enough to worry about identification (provided we don't disaggregate any further) — in the realm of a few dozen people per group — but I might cut the 75-85 bucket and just look at 75+ to be safe.

Now let's start actually looking at stats around time spent on various activities. I still need to resolve the "combined" hours/minutes spent measures.

```{r}
#| label: time-stats
all_data |>
  select(-hgagef, -lscarb, -lscarmn) |>
  filter(hgage >= 0, hgsex >= 0) |>
  # replace negative time-spent values with NA
  mutate(across(starts_with("ls"), ~ case_when(.x < 0 ~ NA, TRUE ~ .x))) |>
  mutate(age_group = cut(hgage, c(15, 25, 35, 45, 55, 65, 75, Inf))) |>
  filter(!is.na(age_group), wave != "a") |>
  select(xwaveid, wave, hgsex, age_group, hgage, everything()) ->
refined_groups

refined_groups |>
  group_by(wave, hgsex, age_group) |>
  summarise(across(starts_with("ls"), list(
    n_obs = ~ length(.x),
    p10 = ~ quantile(.x, 0.1, na.rm = TRUE),
    q1 = ~ quantile(.x, 0.25, na.rm = TRUE),
    mean = ~ mean(.x, na.rm = TRUE),
    q3 = ~ quantile(.x, 0.25, na.rm = TRUE),
    p90 = ~ quantile(.x, 0.9, na.rm = TRUE)
    ))) ->
grouped_stats
```

Now let's visualise the means:

```{r}
#| label: activity-mappings


```

```{r}
#| label: vis-means
#| output: true
survey_vars |>
  pull(long) |>
  set_names(paste0(survey_vars$short, "_mean")) ->
survey_mappings

grouped_stats |>
  select(wave, hgsex, age_group, ends_with("mean")) |>
  pivot_longer(ends_with("mean"), names_to = "activity", values_to = "mean") |>
  mutate(
    sex = case_match(hgsex, 1 ~ "Male", 2 ~ "Female"),
    wave_num = match(wave, letters),
    year = wave_num + 2000,
    activity_long = str_replace_all(activity, survey_mappings)) ->
grouped_stats_recoded

# start with hrs vars
grouped_stats_recoded |>
  filter(str_starts(activity, "lshr")) |>
  ggplot() +
    aes(x = age_group, y = mean, fill = activity_long) +
    geom_col() +
    facet_grid(vars(year), vars(sex)) +
    scale_fill_brewer(type = "qual") +
    theme(
      legend.position = "top",
      legend.direction = "horizontal")
```

This is a good start, and it's already clear that there are big gender discrepancies in terms of tasks like Housework and Playing with your children. It might work better animating the waves, rather than facetting them:

```{r}
#| label: vis-means-animated
#| output: true
grouped_stats_recoded |>
  filter(str_starts(activity, "lshr")) |>
  mutate(activity_long = str_remove(activity_long, "Hours per week - ")) |>
  ggplot() +
    aes(x = age_group, y = mean, fill = activity_long) +
    geom_col() +
    facet_wrap(vars(sex), ncol = 2) +
    transition_time(year) +
    enter_fade() +
    exit_fade() +
    ease_aes("sine-in-out") +
    scale_fill_brewer(type = "qual") +
    theme(
      legend.position = "top",
      legend.direction = "horizontal") +
    labs(
      title = "Hours per week: {floor(frame_time)}"
    ) ->
ani_hrsweek_mean

anim_save("ani_hrsweek_mean.mp4", ani_hrsweek_mean, path = here("out"),
  renderer = ffmpeg_renderer())
```

![Animation: average hours per week spent on unpaid domestic and care work, broken down by activity, age group, sex and year](ani_hrsweek_mean.mp4)
