---
title: Untitled
subtitle: A slightly longer title
format:
  360-analysis-html: default
author: James Goldie
date: last-modified
code-fold: true
---

```{r}
#| label: setup
library(tidyverse)
library(readxl)
library(here)
```

# SDG Indicator 5.4.1

Proportion of time spent on unpaid domestic chores and care work, by sex, age and location (percent)

We'll get the data from the [SDG Global Database](https://unstats.un.org/sdgs/dataportal). They do have a [SDG API](https://unstats.un.org/sdgs/UNSDGAPIV5/swagger/index.html), but I couldn't get the data endpoint to work, so let's work with a manual download instead:

```{r}
#| label: load-data

here("data", "sdg-indicator-5-4-1.xlsx") |>
  read_excel(sheet = "Goal5") |>
  janitor::clean_names() ->
all_data
```

```{r}
#| label: tidy-data
all_data |>
  select(starts_with("series"), geo_area_name, time_period, time_detail, sex, age, location,
    value) |>
  mutate(
    age = fct(age, levels = c(
      "ALLAGE", "3+", "5+", "6+", "6-65",
      "10+", "10-14", "12+", "12-14", "12-24", "14+", "15+", "15-24", "15-49",
      "15-64",  "16+", "18+", "20+", "20-24", "20-64", "20-74", "25-34",
      "25-44", "35-44", "45-54", "45-64", "55-64", "65+", "65-74", "75-84",
      "85+"
    )),
    location = fct(location),
    sex = fct(sex),
    value = as.numeric(value)) ->
all_data_tidy
```

Okay, we have a bit of a mess of age groups. Let's try to see which countries use which age groups:

```{r}
#| label: age-groups

all_data_tidy |>
  select(geo_area_name, age) |>
  group_by(geo_area_name, age) |>
  count() |>
  ungroup() |>
  pivot_wider(names_from = c(age), values_from = c(n), values_fill = 0) |>
  write_csv(here("data", "sdg-age-groups.csv"))
```

Okay, nobody's reporting `ALLAGE`. It looks like most do: 10+, 15+, 15-24, 25-44, 45-54, 55-64, 65+ (there are a few exceptions though). Thankfully, it doesn't appear as if many countries switch age groups based on whether they're breaking down further.

```{r}
#| label: explore
all_data_tidy |>
  filter(age %in% c("10+", "15+", "15-24", "25-44", "45-54", "55-64", "65+")) |>
  filter(geo_area_name == "Australia") |>
  pull(time_detail) |>
  unique()

all_data_tidy |>
  select(geo_area_name, time_detail) |>
  group_by(geo_area_name, time_detail) |>
  count() |>
  ungroup() |>
  pivot_wider(names_from = c(time_detail), values_from = c(n), values_fill = 0) |>
  write_csv(here("data", "sdg-time-groups.csv"))
```

It looks like frequency is all over the place for this data. Let's visualise it more clearly:

```{r}
#| label: calc-frequency
all_data_tidy |>
  select(geo_area_name, time_detail) |>
  group_by(geo_area_name, time_detail) |>
  count() |>
  ungroup() |>
  # average out year ranges
  separate(time_detail, c("start", "end"), sep = "-", fill = "right") |>
  mutate(
    start = as.numeric(start),
    end = as.numeric(end),
    year = if_else(
      is.na(end),
      start,
      start + (end - start) / 2),
    date = date_decimal(year)) |>
  select(-c(start, end, year)) |>
  # sort countries by total number of reports
  group_by(geo_area_name) |>
  add_tally(name = "n_obs") |>
  mutate(country = paste0(geo_area_name, ": ", n_obs)) ->
reporting_frequencies
```

```{r}
#| label: vis-frequencies
ggplot(reporting_frequencies) +
  aes(
    x = fct_reorder(.f = country, .x = n_obs),
    y = date,
    group = geo_area_name) +
  geom_line(colour = "grey") +
  geom_point(fill = "#333333", colour = "white", size = 2, shape = 21) +
  coord_flip() +
  theme_minimal() +
  theme(
    plot.title = element_text(face = "bold"),
    panel.grid.major.y = element_blank()) +
  labs(
    x = NULL, y = NULL,
    title = "How often do countries report on unpaid chores and work?") ->
freq_plot

ggsave(here("out", "assets", "freq-plot.png"), height = 16, width = 9,
  units = "in", bg = "white")
```